---
title: "Do Crows Prefer the City Life or the Peaceful Countryside?"
author: "Andrea Cruz Alegria"
date: "(`r format(Sys.time(), '%d %B %Y')`)"
output: 
  github_document:
    toc: true 
---

# Packages Required
Here are the packages that will be needed throughout the project
```{r, warning=FALSE, message=FALSE}
library(tinytex)
library(rgbif)
library(rnaturalearth)
library(tidyverse)
library(rinat)
library(rnaturalearthdata)
library(raster)
library(viridis)
library(sf)
library(sp)
library(MODIStsp)
library(terra)
library(geodata)
library(xts)
library(rgeoboundaries)
library(rayshader)
library(rgdal)
library(elevatr)
library(remotes)
library(rasterVis)
library(dplyr)
library(ggfortify)
library(devtools)
library(tibble)
library(lubridate)
library(leaflet)
```


To me, crows are one of the most fascinating animals to exist. I am a biologist, so I think that about a lot of animals. But, I think crows are often the most misunderstood creatures. Through history they have been seen as a bad omen, to the point where a group of crows is called a "murder". This has led to some people not taking a liking to crows, and may affect their behavior towards crows, you would think that maybe crows would be driven away by the humans. Nevertheless, crows are quite common wherever you go, they are in the woods and also in populated areas.

How much does human disturbance impact the population of crows? 
In this project we will be looking at the population density in both rural and urban areas to determine which habitat the crows prefer.

![](https://images.unsplash.com/photo-1580420045808-0a7deb24fea7?q=80&w=1933&auto=format&fit=crop&ixlib=rb-4.0.3&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D)


## Research Question
Do Crows prefer the City Life or the Peaceful Countryside? This is the research question we will be investigating.

## Hypothesis
I hypothesize that there would be a higher density in the rural areas compared to urban areas.

# Importation of Data
## Importation of Data from GBIF 

To import data first I will first specify the animal I am interested in, which will be *Corvus corone*. I chose this instead of focusing on all corvids because that would include magpies, rooks, etc, and I just want to focus on the common crow.


```{r, warning=FALSE, message=FALSE}
myspecies <- c("Corvus corone")
```

Now that I've specified the species, I will be using [GBIF](https://www.gbif.org/) as a source to import all *Corvus corone* data (GBIF.org, n.d.).

```{r, warning=FALSE, message=FALSE}

gbif_data <- occ_data(scientificName = myspecies, hasCoordinate = TRUE, limit = 1000, )
occur <- gbif_data$data #naming occurence data 
```

However, I don't want to see all the crows in the world. I want to specifically see those in Switzerland. Thus, I will filiter the data to show *Corvus corone* occurrences in Switzerland.

```{r, warning=FALSE, message=FALSE}

unique(occur$country) #Seeing if Switzerland is indeed listed
gbif_swiss <- occur[occur$country == "Switzerland",]  
(gbif_swiss) #Trying to see first lines of the data

### Renaming Columns of the data 
species <- gbif_swiss$scientificName
Latitude <- gbif_swiss$decimalLatitude
Longitude <- gbif_swiss$decimalLongitude
Date <- gbif_swiss$eventDate

### Creating new data frame with columns of interest 
matrix_gbif <- data.frame(species,     
                          Latitude, 
                          Longitude, 
                          Date)
```

Now we can plot some of this data using the [ggplot](https://ggplot2.tidyverse.org/) function (Wickham, 2016). 

```{r, warning=FALSE, message=FALSE}

## Plots of GBIF
Switzerland <- ne_countries(scale = "medium", returnclass = "sf", country = "Switzerland")


ggplot(data=Switzerland) + 
  geom_sf()  + 
  geom_point(data=matrix_gbif, aes(x = Longitude, y = Latitude), size = 4, shape = 23, fill = "darkgreen") + 
  theme_classic() + 
  labs(title=expression(paste("Occurrence of ",  italic("Corvus corone "), "in Switzerland"))) + 
  xlab("Longitude") + ylab("Latitude")
```


## Importation from iNaturalist 

Because the informaton that I obtained from GBIF is scarce, I will be importing more data from another source called [iNaturalist](https://www.inaturalist.org/) which is a site used by both professionals and amateurs to record their observations.


```{r, warning=FALSE, message=FALSE}

### adding a source column
matrix_gbif$source <- rep("gbif", nrow(matrix_gbif))

### importing data from iNaturalist
inat_swiss <- get_inat_obs(query = "Corvus corone", place_id = "switzerland")

### Renaming Columns of the data 

species_n <- inat_swiss$scientific_name
Latitude_n <- inat_swiss$latitude
Longitude_n <- inat_swiss$longitude
Date_n <- inat_swiss$datetime

### Creating new data frame with columns of interest 
matrix_inat <- data.frame(species_n,     
                          Latitude_n, 
                          Longitude_n, 
                          Date_n)

### renaming the columns in order to merge them together 
colnames(matrix_inat) <- c("species", 
                           "Latitude", 
                           "Longitude", 
                           "Date")

### adding inaturalist as a source
matrix_inat$source <- rep("inat", nrow(matrix_inat))
```

Now, that we've imported the data and cleaned it up a bit, we can merge the two sources into one dataframe. 

```{r, warning=FALSE, message=FALSE}

## Merging the two sources of data together GBIF + iNat
full_matrix <- rbind(matrix_inat, matrix_gbif)
nrow(full_matrix)
```
Now, we are able to create a graph showing the observations of Corvus corone from both sources. We now have a total of 113 observations, which we will work with throughout this paper.


## Plots of GBIF and iNaturalist 
```{r, warning=FALSE, message=FALSE}

ggplot(data = Switzerland) +
  geom_sf()   +
  geom_point(data = full_matrix, aes(x = Longitude, y = Latitude, fill=source), size = 4, 
             shape = 23) + theme_classic() +
  labs(title=expression(paste("Occurrence of Crows in Switzerland"))) + xlab("Longitude") + ylab("Latitude")
```
 This plot allows us to see the crows in relation to switzerland. We can see that there is quite a lot of observations in Ticino. A part from that cluster it seems they are all spread out throughout Switzerland, with no observations in Valais. 
 
 # Elevation Data 
in This section, we will be creating an Elevation map, to see the occurences of *Corvus corone* at different elevations. 
In order to do this, we must first extract the elevation data of Switzerland.

```{r, warning=FALSE, message=FALSE}
#Obtaining elevation data 
elevation_data <- elevatr::get_elev_raster(locations = Switzerland, z = 6, clip = "locations")
points <- SpatialPoints(data.frame(full_matrix$Longitude,full_matrix$Latitude))   
```

## Elevation Plots
Now, we are able to plot a simple map showing the different points where *Corvus corone* are spotted. 

```{r, warning=FALSE, message=FALSE}
#Plots of elevation data 
plot(elevation_data, main=expression(paste("Elevation Points ",  italic("Corvus corone "))), legend.args = list(text="Elevation (m)", side =2, font =2))
plot(points,add=T)

#Extracting the elevation values for each point 
elevation_points <- raster::extract(elevation_data, points, method='bilinear')

#Adding Elevation Column 
full_matrix$Elevation <- elevation_points
```


```{r, warning=FALSE, message=FALSE}
#Checking the data 

mean(full_matrix$Elevation)
```
From this elevation plot, we can see that crows generally tend to be around 1000m, with the highest elevation being 1444m.This means that if they do like rural areas more than urban areas, they do not prefer to be in high altitude mountains. We can further check that the mean, which turns out to be 281m. 

# Temporal Data
Next, we will have a look at what hours *Corvus corone* are being spotted, and during what months they occur the most. 
To do this, we must first add the date and the time to our matrix, and we shall begin with hour occurrence.

```{r, warning=FALSE, message=FALSE}
full_matrix$Date <- ymd_hms(full_matrix$Date)

#Removal of outliers
full_matrix <- subset(full_matrix, full_matrix$Date >= as.Date("2000-01-01"))


#Creation variable with format of date and hour
x_time <- format(full_matrix$Date, format = "%H:%M:%S")   
x_date <- format(full_matrix$Date, format = "%d-%m-%y")  

#Addition of a date column
full_matrix <- add_column(full_matrix, x_date = as.Date(x_date,format = "%d-%m-%y"), .after = 5)

#Addition of a time column
full_matrix <- add_column(full_matrix, Hour = hour(full_matrix$Date), .after = 6)
```

## Temporal Plots
Now that we have added the hour and date to the matrix we can create some plots. 

```{r, warning=FALSE, message=FALSE}
#Plot of corvus corone occurences at different times
ggplot(subset(full_matrix, Hour != 0), aes(x=Hour, colour = source)) +
  geom_histogram(binwidth=1, alpha=1) +
  labs(x = "Hour", y = "Occurrence", title=expression(paste("Occurrence of ", italic("Corvus corone "), "at Different Times ")))+
  scale_x_continuous(breaks = seq(0, 23, by = 1))

```

This plot shows us that most observations occur in the morning at 7:00 a.m. and very little observations after 3:00 p.m., however this is quite biased as it relies on human observations. So, it is not a true reflection of when crows are around.

```{r, warning=FALSE, message=FALSE}
#Addition of columns of each variable 
full_matrix <- add_column(full_matrix, Month = month(full_matrix$Date), .after = 6)
full_matrix <- add_column(full_matrix, Day = day(full_matrix$Date), .after = 7)
full_matrix <- add_column(full_matrix, Year = year(full_matrix$Date), .after = 8)


#Creating a column as a factor to add the labels for each month
full_matrix <- add_column(full_matrix, Month_names = as.factor(cut(full_matrix$Month, 
                                                                   breaks = 12, 
                                                                   labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))), .after = 7)

# Plotting temporal occurrence of corvus coroneover different months
ggplot(full_matrix, aes(x = Month, fill = Month_names)) +
  geom_histogram(bins = 12) +
  scale_x_continuous(breaks = seq(1, 12, by = 1), labels = month.abb) +
  scale_fill_discrete(guide = "none") +
  theme_bw() +
  labs(x = "Month", y = "Frequency", title = expression(paste("Temporal Occurrence of ", italic("Corvus corone "), "in Switzerland "))) +
  theme(strip.text.x = element_text(face = "bold", size = 12))

```
This graph shows us that most observations occur during September-January, and very little between February-May. Most of the observations come from November. This might reflect the breeding times which occur from early April-June, or perhaps people do not go out to observe during the months of February-May. 

```{r, warning=FALSE, message=FALSE}

# Plotting the temporal occurence of Corvus corone at different altitiudes
ggplot(full_matrix, aes(x= Elevation, color = Month_names)) +
  geom_boxplot(alpha = 0.5, linewidth=1) +
  scale_fill_discrete(name = "Month", guide = "legend") +
  theme_bw() +
  theme(plot.title = element_text(size = 12))+
  labs(x = "Elevation", y = "Density", title = expression(paste("Temporal Occurrence of ", italic("Corvus corone "), "in Switzerland ", "at Different Elevations")))
```

We can also combine the temporal occurence and the different elevations. During January they seem to be at higher altitudes than the rest of the months. Other than that, the rest of the year they seem to prefer to be at around 250m.


# Land Cover

We can now start to look at land cover data to see if crows prefer urban or rural areas more. 
We have previously created Switzerland using this code `r Switzerland <- ne_countries(scale = "medium", returnclass = "sf", country = "Switzerland")`. Which is needed here. 

Then we can create a spatial filepath to save the geodata. Then we can use [MODIS](https://modis.gsfc.nasa.gov/data/dataprod/mod12.php) to download the geodata, specifically landcover data MCD12Q1 (Friedl & Sulla-Menashe, 2019). 

The following code is not needed if you have the folder with the images:
```{r, warning=FALSE, message=FALSE}
#defining file path to save downloaded spatial file
# spatial_filepath <- "C:/Users/cruze/OneDrive/RCrow/switzerland_map.shp"
# 
# 
# #Saves spatial_filepath on our computer, but check the path to see if it worked!
# st_write(map_boundary, paste0(spatial_filepath))
# 
# 
# #M*D12Q1 is the code for landcover
# MODIStsp_get_prodlayers("M*D12Q1")
# 
# ### Downloading Landcover data
# MODIStsp(
#   gui = FALSE,
#   out_folder = "C:/Users/cruze/OneDrive/RCrow",
#   out_folder_mod = "C:/Users/cruze/OneDrive/RCrow",
#   selprod = "LandCover_Type_Yearly_500m (MCD12Q1)",
#   bandsel = "LC1",
#   user = "petite.annie14",
#   password = "mqa-qxr6nhp*JWB1zjv",
#   start_date = "2022.01.01",
#   end_date = "2022.12.31",
#   verbose = FALSE,
#   spatmeth = "file",
#   spafile = spatial_filepath,
#   out_format = "GTiff"
# )
```

Now we can extract the landcover data and add it to our dataframe. 

```{r, warning=FALSE, message=FALSE}
#Reading specific landcover data
LC_raster <- raster(here::here("C:/Users/cruze/OneDrive/RCrow/switzerland_map/LandCover_Type_Yearly_500m_v61/LC1/MCD12Q1_LC1_2022_001.tif"))

#Transformation of data, ensuring that coordinates match the rest of the data
LC_raster <- projectRaster(LC_raster, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

#Limiting the data to Switzerland
LC_raster <- raster::mask(LC_raster, as_Spatial(Switzerland ))

#Adding coordinates to the Landcover map
LC <- raster::extract(LC_raster, points, method='simple')
matrix_LC <- as.data.frame(LC_raster, xy = TRUE, na.rm = TRUE) 

# values rounded to the nearest whole number
full_matrix$LC <- round(LC)

# Create character and numeric values (LC and LC_names)
full_matrix$LC_names <- full_matrix$LC 
```

Now that we have the numbers of the LC, we can add the corresponding names using the [MCD12 User Guide](https://lpdaac.usgs.gov/documents/101/MCD12_User_Guide_V6.pdf) using Table 9. For simplicity I put the forets, herbaceous, croplands, and shrublands all under "Vegetated". 

```{r, warning=FALSE, message=FALSE}
### Replace name only in the line LC (in which vector we want to replace the number) [filter]

full_matrix$LC_names[(full_matrix$LC > 0 & full_matrix$LC <= 1)] <- "Barren"
full_matrix$LC_names[(full_matrix$LC > 1 & full_matrix$LC <= 2)] <- "Permanent Snow and Ice"
full_matrix$LC_names[(full_matrix$LC > 2 & full_matrix$LC <= 3)] <- "Water"
full_matrix$LC_names[(full_matrix$LC > 3 & full_matrix$LC <= 9)] <- "Urban"
full_matrix$LC_names[(full_matrix$LC > 9 & full_matrix$LC <= 40)] <- "Vegetated"
full_matrix$LC_names[full_matrix$LC > 40] <- "Unclassified"


custom_colors <- c(
  "Barren" = "orange",
  "Permanent Snow and Ice" = "lightblue",
  "Water" = "blue",
  "Urban" = "red",
  "Vegetated" = "purple",
  "Unclassified" = "grey"
)

full_matrix$LC_colors <- custom_colors[full_matrix$LC_names]
```

## Land Cover Plots

Now that we've colored the land cover types, we can make plots for the land cover to see where crows are observed the most. 


```{r, warning=FALSE, message=FALSE}

opar <- par(no.readonly = TRUE)
par(mar = c(6, 5, 4, 6.5))

plot(LC_raster, legend=FALSE, main=expression(paste("Occurrence of",  italic("Corvus corone "), "in Relation to Land Cover")))
points(points, pch = 20, col = full_matrix$LC_colors)
legend(x = "topright",
       inset = c(-0.5, 0), # You will need to fine-tune the first
       # value depending on the windows size
       legend = names(custom_colors),
       fill = custom_colors, 
       xpd = TRUE) # Needed to put the legend outside the plot

# Back to the default graphical parameters
on.exit(par(opar))

table(full_matrix$LC_names)

```

This graph shows us that observations occur in vegetated and urban areas, and no other areas, however again this is because these observations depend on where humans tread. Visually however, it seems that there are more observations in vegetated areas than in rural areas, even in that big cluster in Ticino. 

We can check the data to see if it correlates to what we can see visually. Indeed it seems that there are 35 urban observations and 78 vegetated (rural) observations. This shows that with the given data set crows do prefer rural areas over urban areas. 


# Other plots 

We have the essential plots, but I would like to dig deeper and find out more. 
This plot shows us the temporal occurrence by land cover. This shows us that most of the observations of crows occur in the morning and the majority in vegetated areas.

```{r, warning=FALSE, message=FALSE}

# Filter for Vegetated and Urban land covers
landcover_subset <- c("Vegetated", "Urban")
subset_matrix <- full_matrix[full_matrix$LC_names %in% landcover_subset, ]

# Plot temporal occurrence for Vegetated and Urban
ggplot(subset_matrix, aes(x = Hour, fill = LC_names)) +
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.7) +
  labs(x = "Hour", y = "Occurrence", title=expression(paste("Temporal Occurrence of",  italic("Corvus corone "), "in Switzerland by Land Cover"))) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal()
```

Furthermore, there is more density of crows in vegetated areas, and also seen throughout the year. In urban areas they have been observed on January, November, October, June, and August.
```{r, warning=FALSE, message=FALSE}
ggplot(full_matrix, aes(x= LC_names, color = Month_names)) +
  geom_boxplot(alpha = 0.5, linewidth=1) +
  scale_fill_discrete(name = "Month", guide = "legend") +
  theme_bw() +
  theme(plot.title = element_text(size = 12))+
  labs(x = "Elevation", y = "Density", title = expression(paste("Temporal Occurrence of ", italic("Corvus corone "), "in Switzerland ", "at Different Elevations")))
```

This graphs allows us to see that indeed, there is more density of crows in vegetated areas than in urban areas. 
```{r, warning=FALSE, message=FALSE}
ggplot(full_matrix, aes(x=LC, fill=LC_names)) + 
    geom_histogram(binwidth=0.6, alpha=0.5) +
    labs(title=expression(paste("Density of ", italic("Corvus corone "), "by Land Cover"))) + xlab("Land Cover Types") + ylab("Density")
```

From this graph we can see that during January and November the crows are seen at higher altitudes in mostly vegetated areas. In October and September we can see more crows are present in urban areas at these times. 

```{r, warning=FALSE, message=FALSE}
ggplot(full_matrix, aes(x = Month, y = Elevation, fill = Month_names)) +
    geom_point(aes(colour = LC_names, alpha=1/20)) +
    scale_x_continuous(breaks = seq(1, 12, by = 1), labels = month.abb) +
    scale_size_continuous(breaks = c(500, 1000, 1500, 2000)) +
    scale_fill_discrete(guide = "none") +
    scale_alpha(guide = "none") +
    labs(title=expression(paste("Activity of ", italic("Corvus corone "), "During the Year at Different Altitudes"))) + xlab("Months") + ylab("Elevation")+
    theme_bw()
```

# Statistics 

Now we can look at some statistic. We will take a look if the choice of land cover is related to season/month, hour, or elevation.

```{r, warning=FALSE, message=FALSE}
# Pearson Correlation
  cor.test(full_matrix$LC,full_matrix$Month)

# Does LC choice depend on season? 
mod1 <- lm(LC~Month,data = full_matrix)
anova(mod1)
```

The statistical tests show there is no significant relationship between the season and land choice.

Is there perhaps a relationship between the moment in the day and the choice of land?

```{r, warning=FALSE, message=FALSE}
# Pearson correlation
cor.test(full_matrix$LC,full_matrix$Hour) 

# Does LC choice depend on the time of day? 
mod2 <- lm(LC~Hour,data = full_matrix)
anova(mod2)
```

There is also no significant relationship between the hour and land choice, with a p-value of 0.16. 


```{r, warning=FALSE, message=FALSE}
# Pearson correlation
cor.test(full_matrix$LC,full_matrix$Elevation) 

# Does LC choice depend on the elevation? 
mod3 <- lm(LC~Elevation,data = full_matrix)
anova(mod3)
```
There is also no significant relationship the land choice and the elevation. 

# Conclusion 
From the graphs that we have created we can see:

- Crows tend to be more active during early morning from 6:30-8:30 a.m.
- They tend to prefer lower elevations ranging from 200-300m, and rarely are they observed at higher altitudes than 400m. 
  + For reference, the Neuchatel train station is at around 400m. 

My research question was looking into whether crows preferred urban or rural areas more. The graphs show that the observations done exclude all land cover types except vegetated and urban. However, we have to keep into account that there's probably no observations being done in permanent snow and ice, barren or in water observations because many people do not go there, and if they do and do see a crow, they might not register it on either GBIF or iNaturalist. 

The plots in the section *Other plots" further confirm that crows are mostly observed in vegetated areas rather than urban areas, there is higher occurence and density in vegetated areas. 

This confirms my hypothesis. However, I must stress that this data might be biased due to it being taken from observations that humans have noted down. This means that the conclusions here should be taken lightly. It might be that there are more crows in urban areas, and people don't note it down on GBIF or iNaturalist, simply because it is common to see crows in those areas. Furthermore, this dataset only contains 113 observations, which may be too little to conclude anything. 

Furthermore the statistical tests show no significant relation between the choice of land by the crows and the hour of the day, the season (month), or elevation. 


# References 
1. Friedl, M., & Sulla-Menashe, D. (2019). MCD12Q1 MODIS/Terra+Aqua Land Cover Type Yearly L3 Global 500m SIN Grid V006 [dataset]. NASA EOSDIS Land Processes Distributed Active Archive Center. https://doi.org/10.5067/MODIS/MCD12Q1.006
2. GBIF.org. (n.d.). GBIF Occurence Download [dataset]. https://doi.org/10.15468/dl.0eksga
3. Wickham, H. (2016). ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org


